---
title: "Tobias Schatton - Diffbind treatment"
output:
   html_document:
      code_folding: hide
      df_print: paged
      highlights: pygments
      number_sections: true
      self_contained: true
      theme: default
      toc: true
      toc_float:
         collapsed: true
         smooth_scroll: true
date: "2023-08-31"
---

## Set-up

Here we load packages and the data ahead of our analyses.
```{r setup, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(dev = c('png', 'cairo_pdf'),
                      fig.align = 'center', 
                      fig.height = 5, 
                      fig.width = 7,
                      pdf.options(encoding = "ISOLatin9.enc"),
                      fig.path='figures/',
                      warning=FALSE, 
                      message=FALSE,
                      cache = FALSE,
                      dev = c("png", "pdf"),
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE, echo=FALSE)
```

```{r load-libraries}
# Load libraries
library(fgsea)
library(clusterProfiler)
library(SummarizedExperiment)
library(gridExtra)
library(DESeq2)
library(tidyverse)
library(ggrepel)
library(knitr)
library(pheatmap)
library(RColorBrewer)
library(apeglm)
library(Category)
library(ChIPpeakAnno)
library(DEGreport)
library(DiffBind)
library(ChIPseeker)
library(csaw)
library(Vennerable)
library(reshape2)
library(org.Mm.eg.db)
library(ReactomePA)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene



# Set ggplot2 default theme
ggplot2::theme_set(theme_light(base_size = 14))

sanitize_datatable = function(df, ...) {
 # remove dashes which cause wrapping
 DT::datatable(df, ..., rownames=gsub("-", "_", rownames(df)),
                   colnames=gsub("-", "_", colnames(df)))
}


pruned_annotations <- read.csv("biomart_entrez.csv") %>% .[,-1]

pruned_annotations2 = pruned_annotations %>%
    arrange(entrezgene_id, ensembl_gene_id) %>%
    distinct(entrezgene_id, .keep_all = TRUE)

```



## Count Reads

We load in results from Diffbind Count
```{r}
## Load from O2 run


all <- readRDS("//n/data1/cores/bcbio/PIs/tobias_schatton/schatton_atacseq_hbc04813/diffbind/bowtie2/all_counts.RDS")




```

## Change paths and add info

```{r}


all$class[4, ]<- c(rep("alpha",2), rep("beta",2), rep("untreated",2))

```

## Normalize and manually add back in metadata and contrasts

We normalize our counts and add back in metadata for our DE run
```{r}
## alpha vs. beta

all_norm1 <- dba.normalize(all)

## Extract Normalized counts for later

normCounts_all <- dba.peakset(all_norm1, bRetrieve=TRUE, DataType=DBA_DATA_FRAME)

write.csv(normCounts_all, file="counts_all_samples.csv")

```


## Try to make better PCA plots

We extracted normalized counts and re-do the PCA plots ourselves.

```{r}

meta_all <- data.frame(colnames(normCounts_all)[5:10],c("Alpha","Alpha","Beta","Beta","Untreated","Untreated"))
colnames(meta_all) <- c("sample","group")
rownames(meta_all) <- meta_all$sample

log_norm_all <- log2(normCounts_all[,5:10]+1)

pca <- degPCA(log_norm_all, meta_all, condition = "group", name = "sample", data = T)[["plot"]]

pca_labels <- pca[["labels"]]
pca_data <- pca[["data"]] %>% as_tibble()
pca_data %>% 
    ggplot(aes(x = PC1, y = PC2, color = group)) + 
    geom_point(size=4) +  
    xlab(pca_labels$x) + 
    ylab(pca_labels$y) +
    ggtitle(paste0("All samples", "\nPCA using ", nrow(normCounts_all), " peaks")) +
    theme(plot.title=element_text(hjust=0.5)) + scale_color_manual(values = c("firebrick1","firebrick4","grey58")) 


pdf("PCA_plot_all.pdf", width = 6, height = 5)

pca_data %>% 
    ggplot(aes(x = PC1, y = PC2, color = group)) + 
    geom_point(size=5) +  
    xlab(pca_labels$x) + 
    ylab(pca_labels$y) +
    ggtitle(paste0("All samples", "\nPCA using ", nrow(normCounts_all), " peaks")) +
    theme(plot.title=element_text(hjust=0.5)) + scale_color_manual(values = c("firebrick1","firebrick4","grey58")) 

dev.off()

```

## Get DB peaks

```{r}
all_db <- readRDS("/n/data1/cores/bcbio/PIs/tobias_schatton/schatton_atacseq_hbc04813/diffbind/bowtie2/all_counts_dbpeaks.RDS")

all_db$class[4, ]<- c(rep("alpha",2), rep("beta",2), rep("untreated",2))

## Extract Normalized counts for later

all_norm_db <- dba.peakset(all_db) 

normCounts_all_db <- dba.peakset(all_norm_db, bRetrieve=TRUE, DataType=DBA_DATA_FRAME)

write.csv(normCounts_all_db, file="counts_all_samples_dbpeaks.csv")

```


## Make heatmap
```{r, height=25, width=10}
## We have waaay too many peaks to plot. We randomly sub-sample 3k

ar_it_rand_idx <- sample(34577,3000, replace=FALSE)

normCounts_all_db_subset <- normCounts_all_db[ar_it_rand_idx,]

meta <- c(rep("IFAlpha",2), rep("IFBeta",2), rep("Untreated",2)) %>% data.frame()
colnames(meta) <- "Groups"
rownames(meta) <- colnames(normCounts_all_db_subset)[4:9]
### Set a color palette
heat_colors <- colorRampPalette(c("blue", "black", "yellow"))(100)
anno_colors <- list(Groups=c(IFAlpha="firebrick1",IFBeta="firebrick4", Untreated="grey58"))


### Run pheatmap using the metadata data frame for the annotation (11 x 5)

pdf("heatmap.pdf", height=20, width=10)
pheatmap(normCounts_all_db_subset[,4:9], 
    color = heat_colors, 
    cluster_rows = T, 
    show_rownames = F,
    annotation = meta, 
    annotation_colors = anno_colors,
    border_color = NA, 
    fontsize = 10, 
    scale = "row", 
    fontsize_row = 10, 
    height = 20)

dev.off()
```

# R session

List and version of tools used for the report.

```{r}
sessionInfo()
```